# Interactive component

```{r }
library(readr)
library(vcd)
library(ggalluvial)
library(dplyr)
library(plotly)
library(d3r)
library(tidyverse)
library(leaflet)
library(formattable)
```

Formula 1 is a sport where teams and drivers strive to exceed their performance. It is therefore important to analyze the time for each race and each lap in order to look for details to improve. Moreover, its performance compared to the competitors is a major asset to win races. 

The graph opposite allows to compare the times between two drivers on the same race or on the same race but in different years for the same driver. 

You just have to choose the year, the race and the drivers and then after pressing "GO" the graph is displayed.  


```{r,echo=FALSE, results="asis"}
library(readr)
library("jsonlite")

circuits <- read_csv("data/circuits.csv")
constructor_results <- read_csv("data/constructor_results.csv")
constructor_standings <- read_csv("data/constructor_standings.csv")
constructors <- read_csv("data/constructors.csv")
driver_standings <- read_csv("data/driver_standings.csv")
drivers <- read_csv("data/drivers.csv")
lap_times <- read_csv("data/lap_times.csv")
pit_stops <- read_csv("data/pit_stops.csv")
qualifying <- read_csv("data/qualifying.csv")
races <- read_csv("data/races.csv")
results <- read_csv("data/results.csv")
status <- read_csv("data/status.csv")

cat(
  paste(
  '<script>
    var drivers = ',toJSON(drivers),';
  </script>'
  , sep="")
)

cat(
  paste(
  '<script>
    var races = ',toJSON(races),';
  </script>'
  , sep="")
)

cat(
  paste(
  '<script>
    var results = ',toJSON(results),';
  </script>'
  , sep="")
)

cat(
  paste(
  '<script>
    var lap_times = ',toJSON(lap_times),';
  </script>'
  , sep="")
)
```

<html lang="en-US">
<meta charset="utf-8">
<head>
  <!-- Load in the d3 library -->
  <script src="https://d3js.org/d3.v7.js"></script>
  
<style type ="text/css">
      .linestyle1 {
        fill: none;
        stroke: red;
        stroke-width: 2px;
      }
      .linestyle2 {
        fill: none;
        stroke: teal;
        stroke-width: 2px;
      }
      
</style>
</head>

<body>
<select id="selectButtonyear"></select>
<select id="selectButtonrace"></select>
<select id="selectButtondriver1"></select>
<select id="selectButtondriver2"></select>
<button type="button" id="selectButtongo">Go</button>

<div id="my_dataviz"></div>
  <svg align = center width="160" height="800"> 
  </svg> 

<script>

// Button Year
var years = [];
var i = 1;
var l = 2021;
for (var i = 1950; i < l + 1; i++) {
    years.push(i);
}

var years = years.reverse()

d3.select("#selectButtonyear")
      .selectAll('myOptions')
     	.data(years)
      .enter()
    	.append('option')
      .text(function (d) { return d; }) // text showed in the menu
      .attr("value", function (d) { return d; })


d3.select("#selectButtonyear").on("change", function(d) {
        var v1 = document.getElementById("selectButtonyear").value
        var race = []
        var l = races["length"]
        var existingraces = []
        var llp = lap_times["length"]
        for (var i = 1; i < llp; i++) {
  existingraces.push(lap_times[i]["raceId"])
  }
        var uniqueArr = [...new Set(existingraces)]
        for (var i = 1; i < l; i++) {
  if(races[i]["year"] == v1){
  if(uniqueArr.includes(races[i]["raceId"])){
  race.push(races[i]["name"])
  
  }}
}
      d3.select("#selectButtonrace")
      .selectAll('myOptions')
     	.data(race)
      .enter()
    	.append('option')
      .text(function (d) { return d; }) // text showed in the menu
      .attr("value", function (d) { return d; })
      
      d3.select("#selectButtonrace").on("change", function(d) {
        var v1 = document.getElementById("selectButtonyear").value
        var v2 = document.getElementById("selectButtonrace").value
        var raceid = races.filter(obj=> obj.year == v1).filter(obj=> obj.name == v2)[0]["raceId"];
        var pdriver = lap_times.filter(obj=> obj.raceId == raceid);
        var existingdrivers = []
        var led = pdriver["length"]
        for (var i = 1; i < led; i++) {
  existingdrivers.push(pdriver[i]["driverId"])
  }
        var uniqueArrd = [...new Set(existingdrivers)]
        var d = results.filter(obj=> obj.raceId == raceid)
        var driver = []
        var ld = d["length"]
        for (var i = 1; i < ld ; i++) {
       
    driver.push(d[i]["driverId"]);
}

drivername = drivers.filter(function(obj) {return uniqueArrd.includes(obj.driverId);})
var drivernamelist = []
var ldn = drivername["length"]
for (var i = 0; i < ldn ; i++) {
    drivernamelist.push(drivername[i]["surname"]);
}

        
      d3.select("#selectButtondriver1")
      .selectAll('myOptions')
     	.data(drivernamelist)
      .enter()
    	.append('option')
      .text(function (d) { return d; }) // text showed in the menu
      .attr("value", function (d) { return d; })
      
       d3.select("#selectButtondriver2")
      .selectAll('myOptions')
     	.data(drivernamelist)
      .enter()
    	.append('option')
      .text(function (d) { return d; }) // text showed in the menu
      .attr("value", function (d) { return d; })
      
d3.select("#selectButtongo").on("click", function(d) {

var driver1name = document.getElementById("selectButtondriver1").value
var driver1id = drivers.filter(obj=> obj.surname == driver1name)[0]["driverId"]
var driver1laps = lap_times.filter(obj=> obj.raceId == raceid).filter(obj=> obj.driverId == driver1id)
      
var ldriver1laps = driver1laps["length"]
      
var lap_time1 = []
      
for (var i = 0; i < ldriver1laps; i++) {
    lap_time1.push([driver1laps[i]["lap"],driver1laps[i]["milliseconds"]]);
}

var time1 = []
var ltime1 = lap_time1.length

for (var i = 0; i < ltime1; i++) {
    time1.push(lap_time1[i][1]);
}

var driver2name = document.getElementById("selectButtondriver2").value
var driver2id = drivers.filter(obj=> obj.surname == driver2name)[0]["driverId"]
var driver2laps = lap_times.filter(obj=> obj.raceId == raceid).filter(obj=> obj.driverId == driver2id)
      
var ldriver2laps = driver2laps["length"]
      
var lap_time2 = []
      
for (var i = 0; i < ldriver2laps; i++) {
    lap_time2.push([driver2laps[i]["lap"],driver2laps[i]["milliseconds"]]);
}

var time2 = []
var ltime2 = lap_time2.length

for (var i = 0; i < ltime2; i++) {
    time2.push(lap_time2[i][1]);
}

var min = Math.min(...time1)
var max = Math.max(...time1)
var mean = time1.reduce((a,b) => a + b, 0) / time1.length
var nlaps = lap_time1.length

var data1 = driver1laps;
var data2 = driver2laps;


var margin = {top: 70, right: 100, bottom: 50, left: 70},
    width = 850 - margin.left - margin.right,
    height = 400 - margin.top - margin.bottom;
    
var svg = d3.select("#my_dataviz")
  .append("svg")
    .attr("width", width + margin.left + margin.right)
    .attr("height", height + margin.top + margin.bottom)
  .append("g")
    .attr("transform",
          "translate(" + margin.left + "," + margin.top + ")");

var x = d3.scaleLinear()
    .domain([0, nlaps + 1])
    .range([ 0, width ]);
  svg.append("g")
    .attr("transform", "translate(0," + height + ")")
    .call(d3.axisBottom(x));

svg.append("text")             
      .attr("transform",
            "translate(" + (width/2) + " ," + 
                           (height + margin.top - 40) + ")")
      .style("text-anchor", "middle")
      .text("Laps");
    
var y = d3.scaleLinear()
    .domain([min-5000, mean + 10000])
    .range([ height, 0]);
  svg.append("g")
    .call(d3.axisLeft(y));
    
svg.append("text")
      .attr("transform", "rotate(-90)")
      .attr("y", 0 - margin.left - 3)
      .attr("x",0 - (height / 2))
      .attr("dy", "1em")
      .style("text-anchor", "middle")
      .text("Time (milliseconds)");  

var tooltip = d3.select("#my_dataviz")
    .append("div")
    .style("opacity", 0)
    .attr("class", "tooltip")
    .style("background-color", "white")
    .style("border", "solid")
    .style("border-width", "1px")
    .style("border-radius", "5px")
    .style("padding", "10px")


     
 svg.append('g')
    .selectAll("dot")
    .data(data1)
    .enter()
    .append("circle")
      .attr("class", "blue dot")
      .attr("cx", function (d) { return x(d.lap); } )
      .attr("cy", function (d) { return y(d.milliseconds); } )
      .attr("r", 3)
      .style("fill", "#69b3a2")
    .on('mouseover', function (d) {
     d3.select(this).transition()
          .duration('10')
          .attr("r", 7)
       tooltip
          .style("opacity", 1)
          .html("Lap : " + d.lap);})
      .on('mouseout', function (d) {
     d3.select(this).transition()
          .duration('20')
          .attr("r", 3)
          tooltip
      .transition()
      .duration(20)
      .style("opacity", 0);});
      
 svg.append('g')
    .selectAll("dot")
    .data(data2)
    .enter()
    .append("circle")
      .attr("class", "red dot")
      .attr("cx", function (d) { return x(d.lap); } )
      .attr("cy", function (d) { return y(d.milliseconds); } )
      .attr("r", 3)
      .style("fill", "#d66074")
      .on('mouseover', function (d) {
     d3.select(this).transition()
          .duration('10')
          .attr("r", 7)
       tooltip
          .style("opacity", 1)
          .html("Time : " + d.milliseconds);})
      .on('mouseout', function (d, i) {
     d3.select(this).transition()
          .duration('20')
          .attr("r", 3)
        tooltip
      .transition()
      .duration(20)
      .style("opacity", 0);});
      

var dotdriver1 = [{x: nlaps - 13, y : mean + 8000, label : driver1name}]
var dotdriver2 = [{"x": nlaps - 13, "y" : mean + 6000, "label" : driver2name}]


 svg.append('g')
    .selectAll("dot")
    .data(dotdriver1)
    .enter()
    .append("circle")
      .attr("class", "dotlabeldriver1")
      .attr("cx", function (d) { return x(d.x); } )
      .attr("cy", function (d) { return y(d.y); } )
      .attr("r", 3)
      .style("fill", "#69b3a2");
    
      
 svg.append('g')
    .selectAll("dot")
    .data(dotdriver1)
    .enter()
    .append('g')
    .append("text")
            .text(function(d){
                    return d.label;
                })
            .attr("x", function (d) {
                    return x(d.x + 1);
                })
            .attr("y", function (d) {
                    return y(d.y);
                });
      

 svg.append('g')
    .selectAll("dot")
    .data(dotdriver2)
    .enter()
    .append("circle")
      .attr("class", "dotlabeldriver2")
      .attr("cx", function (d) { return x(d.x); } )
      .attr("cy", function (d) { return y(d.y); } )
      .attr("r", 3)
      .style("fill", "#d66074");
    
      
 svg.append('g')
    .selectAll("dot")
    .data(dotdriver2)
    .enter()
    .append('g')
    .append("text")
            .text(function(d){
                    return d.label;
                })
            .attr("x", function (d) {
                    return x(d.x + 1);
                })
            .attr("y", function (d) {
                    return y(d.y);
                });
})
      

    })
      
    })
    


      
    
</script>
</body>
</html>

